---
output: html_document
---
<link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css">
<style> 
  .right{ 
    float: right; 
  } 
  
  img{ 
    display:block; 
    margin: 0 auto;
  } 
  
  pre{ 
    overflow-x: hidden; 
    overflow-y: hidden; 
  } 
  
  body{ 
    margin-top: 20px; 
    margin-bottom: 20px; 
  } 
  
  .col-left{ 
    float: left; 
    width: 50%; 
  } 
  
  .col-right{ 
    float: right; 
    width: 50%; 
  } 
  
  p{
    margin-bottom: 0px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
```

<div class='jumbotron'>
  <h2 class='display-3 text-uppercase'>Week 1 Assignment</h2>
  <h4 class='right text-uppercase'>By Brian Weinfeld</h4>
  <div class='clearfix'></div>
  <h5 class='right text-uppercase'>January 30, 2018</h5>
</div>

<div class='alert alert-info'>
  I began by reading 2 in csv files from my github page. The first contains the unaltered raw mushroom data while the second contains a lightly hand-edited version of the attribute names, abbreviations and associated values to be used with the mushroom data.
</div>

```{r}
mushrooms_raw <- read.csv('https://raw.githubusercontent.com/brian-cuny/607assignment1/master/agaricus-lepiota.csv', header=FALSE, stringsAsFactors=FALSE)
names <- read.csv('https://raw.githubusercontent.com/brian-cuny/607assignment1/master/names.csv', header=FALSE, stringsAsFactors=FALSE, row.names=1, col.names=0:12)
```

<div class='alert alert-info'>
  I transformed names dataframe for easier processing and then used the column names to assign the columns of mushrooms_raw.
</div>

```{r}
names <- t(names)
names(mushrooms_raw) <- colnames(names)
head(mushrooms_raw[,1:5], 10)
head(names[,1:5], 10)
```

<div class='alert alert-info'>
  I subset the raw mushroom data by selected 'classes' and 3 other columns ('odor', 'population' and 'habitat'). I stored this subset in mushrooms_reduced and stored the columns in my_columns for future use.
</div>

```{r}
my_columns <- c('classes', 'odor', 'population', 'habitat')
mushrooms_reduced <- subset(mushrooms_raw, select=my_columns)
head(mushrooms_reduced, 10)
```

<div class='alert alert-info'>
  The below method accepts the names and mushroom_reduced data structures and the name of the column that I wish to process. It reads the names matrix and creates two vectors that contain the matching abbrevation and full names of the given class. An a to b substitution is then made and the modified data is returned.
</div>

```{r}
#"unlist" code adapted from "Henry" on Stackoverflow
process_names <- function(names, mushrooms_reduced, column){
  split_names <- strsplit(names[,column], '=')
  split_names[sapply(split_names, length)==0] <- NULL
  abbreviations <- unlist(split_names)[2*(1:length(split_names))]
  full <- unlist(split_names)[2*(1:length(split_names))-1]
  return(mapvalues(mushrooms_reduced[,column], from=abbreviations, to=full))
}
```

<div class='alert alert-info'>
  I looped over all the appropriate columns calling the process_names method and this will substitute the abbrevation for the full name in every entry in the data frame.
</div>

```{r}
for(column in my_columns){
  mushrooms_reduced[,column] <- process_names(names, mushrooms_reduced, column)
}
head(mushrooms_reduced, 10)
```

<div class='alert alert-success'>
  In conclusion, the original mushrooms data has been read in, subset and updated with more readable information and is ready to be saved or processed further.
</div>






